%% ============================================================
%% Applying OAuth 2.1 Token Exchange with Nested Actor Claims
%% Approximate word count (excluding abstract, references, captions): ~8500
%% to MCP Agentic Workflows
%% v6 — Auth0 real-IdP validation: new subsection + Table VI, abstract updated,
%%       Future Work item 3 completed and removed, Conclusion bullet 5 added
%%
%% Compile: pdflatex (run twice for cross-references)
%% Packages: IEEEtran (journal), tikz, pgfplots,
%%           booktabs, multirow, colortbl, tabularx,
%%           amsmath (multline)
%% ============================================================
\documentclass[journal]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{url}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{multirow}
\usepackage{array}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{colortbl}
\usepackage{tabularx}
\usepackage{balance}
\usepackage{placeins}   %% \FloatBarrier — prevents floats crossing section boundaries

\usetikzlibrary{
  arrows.meta, shapes, positioning, fit,
  backgrounds, calc, decorations.pathreplacing,
  decorations.text
}
\pgfplotsset{compat=1.18}
\usepgfplotslibrary{groupplots}

\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}

%% ── colour palette ──────────────────────────────────────────
\definecolor{C_BASE}{HTML}{E07B54}  % warm orange – insecure / baseline
\definecolor{C_SEC} {HTML}{4C9BE8}  % cool blue   – NAC / secure
\definecolor{C_PAR} {HTML}{5DBB7A}  % green       – blocked / parallel
\definecolor{C_WARN}{HTML}{E74C3C}  % red         – attack / warning
\definecolor{C_HEAD}{HTML}{2C3E50}  % dark navy   – headers
\definecolor{SD_BG} {HTML}{EBF4FF}  % light blue  – sequence actor bg

%% ── code listing style ──────────────────────────────────────
\lstset{
  basicstyle=\scriptsize\ttfamily,
  breaklines=true,
  frame=single,
  captionpos=b,
  xleftmargin=1.2em,
  numbers=left,
  numberstyle=\tiny\color{gray},
  numbersep=4pt,
}

%% ── global TikZ styles ──────────────────────────────────────
\tikzset{
  %% -- sequence diagram actors --
  SD/actor/.style={
    draw=C_HEAD!60, rounded corners=3pt,
    fill=SD_BG, minimum width=2.3cm, minimum height=0.58cm,
    align=center, font=\scriptsize\bfseries
  },
  %% -- sequence diagram lifeline --
  SD/life/.style={draw, dashed, gray!55, thin},
  %% -- sequence diagram forward arrow --
  SD/fwd/.style={-{Stealth[length=3.5pt]}, thick, color=C_HEAD!80},
  %% -- sequence diagram response arrow --
  SD/bck/.style={-{Stealth[length=3.5pt]}, thick, dashed, color=gray!65},
  %% -- sequence diagram message label --
  SD/lbl/.style={font=\tiny, fill=white, inner sep=1.5pt, align=center},
  %% -- sequence diagram "par" combined fragment --
  SD/par/.style={draw=gray, dashed, thick, rounded corners=2pt},
  %% -- generic box (architecture diagrams) --
  arch/hub/.style={
    draw, rounded corners=4pt, fill=C_SEC!18,
    minimum width=1.9cm, minimum height=0.6cm,
    align=center, font=\scriptsize\bfseries
  },
  arch/worker/.style={
    draw, rounded corners=4pt, fill=C_PAR!22,
    minimum width=1.6cm, minimum height=0.55cm,
    align=center, font=\scriptsize
  },
  arch/oauth/.style={
    draw, rounded corners=4pt, fill=C_BASE!18,
    minimum width=1.9cm, minimum height=0.6cm,
    align=center, font=\scriptsize\bfseries
  },
  arch/user/.style={
    draw, rounded corners=4pt, fill=gray!14,
    minimum width=1.4cm, minimum height=0.55cm,
    align=center, font=\scriptsize\bfseries
  },
  arch/arr/.style={-{Stealth[length=3.5pt]}, thick},
  arch/darr/.style={-{Stealth[length=3.5pt]}, thick, dashed, gray!70},
}

%% ============================================================
\begin{document}

\title{Applying OAuth 2.1 Token Exchange with Nested Actor Claims\\
to MCP Agentic Workflows:\\
An Architectural Pattern for Secure Delegation}

\author{\IEEEauthorblockN{Naitik Patel}
\IEEEauthorblockA{Independent Researcher\\
naitikpatel044@gmail.com}}

\maketitle

%% ============================================================
\begin{abstract}
The Model Context Protocol (MCP) has emerged as the standard interface for
connecting large language model agents to downstream tool services. A pervasive
security anti-pattern---\textit{token passthrough}---causes orchestrating hubs
to forward the user's OAuth access token unchanged to every downstream service,
simultaneously violating audience binding, scope attenuation, and delegation
chain visibility, constituting the \textit{Confused Deputy Problem}. We propose
an architectural pattern based on RFC~8693 (OAuth~2.0 Token Exchange) with
Nested Actor Claims (NAC), in which each delegation hop issues a fresh,
narrowly scoped child token valid only for the target service. We construct a
reference implementation and evaluate it against four concrete attack vectors
using an ablation study across multiple enforcement configurations, validated
against a live identity provider. The evaluation confirms that the pattern
eliminates all identified attack vectors, that each of the three security
properties is independently necessary, and that per-hop overhead scales
linearly with chain depth, adding only modest overhead that is projected to
remain in the single-digit to low-double-digit percentage range in typical
production deployments. The pattern is authorization-server-agnostic and
requires no changes to the agent layer.
\end{abstract}

\begin{IEEEkeywords}
Model Context Protocol, OAuth 2.1, RFC 8693, Token Exchange, Nested Actor Claims,
Confused Deputy, Multi-Agent AI, Delegation Chain, Agentic Security, OWASP GenAI
\end{IEEEkeywords}

%% Plain Language Summary (for broad audience accessibility)
\vspace{4pt}
\noindent\textit{\textbf{Plain Language Summary:}}
\textit{AI assistants increasingly call multiple services on a user's behalf,
carrying a security credential called an OAuth token. A common
shortcut---passing the same token to every service unchanged---is insecure: it
gives each service far more access than it needs, enables attackers to misuse the
token across service boundaries, and leaves no audit trail showing which AI agent
made each call. We describe and evaluate a standards-based fix: before calling
each service, the AI hub asks a trusted authorization server for a fresh,
narrowly scoped token valid only for that service. Our evaluation shows this
blocks all four known attacks with no false positives, adds approximately 2\,ms
per service call in raw cryptographic cost ($\approx$35\% single-machine overhead;
projected $+7$--$15\%$ in production with a dedicated authorization server),
and leaves a complete, verifiable record of every delegation step.}

%% ============================================================
\section{Introduction}
\label{sec:introduction}
%% ============================================================

Autonomous AI systems are rapidly transitioning from single-model assistants to
distributed multi-agent orchestrators. A user asks an AI assistant to
``review my calendar, summarise the relevant notes, send the summary by email, and
post an update to the team channel.'' Behind this single natural-language request
lies a coordinated sequence of calls to four distinct downstream services, each
with its own authorization requirements.

The Model Context Protocol (MCP) \cite{mcp_spec,mcp_wiki}, introduced by Anthropic in
November 2024 and subsequently adopted across the industry, provides the emerging
standard interface for this hub-and-spoke architecture: an \textit{assistant hub}
receives tool calls from an agent and in turn calls \textit{worker services} as
an HTTP client. This intermediary role creates an authorization challenge. The
user delegates authority to the hub by providing an OAuth access token. The hub
must somehow convey appropriate authorization to each downstream worker.

The natural, default approach is \textbf{token passthrough}: the hub forwards
the original user token unchanged to every service it calls. Services respond,
and workflows complete---so the problem is invisible at runtime. However,
token passthrough is deeply insecure. The MCP security specification \cite{mcp_security} explicitly
identifies token passthrough as an \textit{anti-pattern} that ``creates confused
deputy risks,'' and the Coalition for Secure AI (CoSAI) whitepaper on MCP
security \cite{cosai_mcp} states: ``don't pass through OAuth tokens directly.
Use token exchange (RFC 8693) to maintain full accountability and prevent confused
deputy attacks.''

Despite this consensus recommendation, no prior work has (1)~formally
characterised the three security properties that token passthrough
violates \textit{audience binding}, \textit{scope attenuation}, and
\textit{delegation chain visibility} (2)~described an architectural pattern for
remediation that is implementation-agnostic, (3)~provided an ablation study
demonstrating the independent necessity of each property, or (4)~empirically
quantified the overhead of the remediated pattern. This paper addresses all four
gaps.

\textbf{Contributions:}
\begin{enumerate}
  \item \textbf{Formal problem characterisation.} We define three independent
        security properties (P1 audience binding, P2 scope attenuation, P3
        delegation chain visibility) and prove that token passthrough violates
        all three simultaneously (Section~\ref{sec:problem}).
  \item \textbf{Formal solution with proofs.} We propose the RFC~8693 NAC
        architectural pattern and prove, by construction, that it satisfies
        P1--P3 and prevents token replay (Theorems~1--4,
        Section~\ref{sec:solution}).
  \item \textbf{Theoretical overhead bound.} We prove that per-hop overhead is
        $O(k)$ sequentially and $O(1)$ in wall-clock time under concurrent
        issuance, with an irreducible constant of one RSA signing operation
        plus one revocation store write (Proposition~1,
        Section~\ref{sec:solution}).
  \item \textbf{Empirical validation.} A reference implementation and 720-trial
        ablation study confirm all four theorems and the overhead bound, with
        zero false positives and $R^2 = 0.9999$ linearity (Section~\ref{sec:eval}).
\end{enumerate}

%% ============================================================
\section{Background and Definitions}
\label{sec:background}
%% ============================================================

This section establishes the technical context for the paper.

\subsection{LLM Agents and the ReAct Model}

Modern LLMs exhibit emergent tool-use capabilities. In the \textit{ReAct}
paradigm \cite{yao2022react}, the model interleaves reasoning with action
calls: it selects a tool, calls it, observes the result, and re-plans. This
is the computational model underlying MCP tool use \cite{wang2024survey}.

\subsection{The Model Context Protocol}

MCP \cite{mcp_spec} defines a client-server protocol over JSON-RPC 2.0. An
\textit{MCP server} exposes \textit{tools} (callable functions), \textit{resources}
(data), and \textit{prompts} (templates) to an \textit{MCP client}. Authentication
is carried via the \texttt{Authorization: Bearer} HTTP header.

The MCP authorization specification \cite{mcp_spec_2025} treats MCP servers as
OAuth 2.1 resource servers. Figure~\ref{fig:mcp_std_auth} shows the standard
authorization flow: the MCP client discovers the authorization server via
Protected Resource Metadata (PRM), performs an OAuth 2.1 authorization code flow
with PKCE, obtains an access token, and uses it to call the MCP server. This flow
governs the \textit{client-to-hub} boundary only. It does not specify how the hub
should authorize its own calls to downstream worker services---the gap this paper
fills.

\begin{figure*}[!ht]
\centering
\begin{tikzpicture}[
  every node/.style={font=\scriptsize},
  x=1cm, y=1cm,
]
%% ── Actor x-positions ────────────────────────────────────────
%% textwidth ≈ 17cm; 4 actors, each 2.4cm wide, centred as below
\pgfmathsetmacro{\xC}{1.8}   % MCP Client
\pgfmathsetmacro{\xS}{6.8}   % MCP Server
\pgfmathsetmacro{\xP}{11.8}  % PRM
\pgfmathsetmacro{\xA}{16.3}  % Auth Server

%% ── Top actor boxes ──────────────────────────────────────────
\node[SD/actor, minimum width=2.3cm] (C1) at (\xC, 0)
  {MCP Client\\(Agent)};
\node[SD/actor, minimum width=2.3cm] (S1) at (\xS, 0)
  {MCP Server};
\node[SD/actor, minimum width=2.8cm] (P1) at (\xP, 0)
  {Protected Resource\\Metadata (PRM)};
\node[SD/actor, minimum width=2.3cm] (A1) at (\xA, 0)
  {Authorization\\Server};

%% ── Lifelines ────────────────────────────────────────────────
\draw[SD/life] (\xC,-0.32) -- (\xC,-11.6);
\draw[SD/life] (\xS,-0.32) -- (\xS,-11.6);
\draw[SD/life] (\xP,-0.32) -- (\xP,-11.6);
\draw[SD/life] (\xA,-0.32) -- (\xA,-11.6);

%% ── Messages ─────────────────────────────────────────────────
%% 1. Client → Server  (solid)
\draw[SD/fwd] (\xC,-1.2) -- (\xS,-1.2);
\node[SD/lbl, above] at ({(\xC+\xS)/2},-1.2)
  {1.\ Request to call tool (e.g., \texttt{tools/call})};

%% 2. Server → Client  (dashed, 401)
\draw[SD/bck] (\xS,-2.2) -- (\xC,-2.2);
\node[SD/lbl, above] at ({(\xC+\xS)/2},-2.2)
  {2.\ 401 Unauthorized with PRM URI (\texttt{WWW-Authenticate} header)};

%% 3. Client → PRM  (solid)
\draw[SD/fwd] (\xC,-3.2) -- (\xP,-3.2);
\node[SD/lbl, above] at ({(\xC+\xP)/2},-3.2)
  {3.\ Fetch PRM document};

%% 4. PRM → Client  (dashed)
\draw[SD/bck] (\xP,-4.2) -- (\xC,-4.2);
\node[SD/lbl, above] at ({(\xC+\xP)/2},-4.2)
  {4.\ Return metadata (authz server, scopes, token types)};

%% 5. Client → Auth  (solid)
\draw[SD/fwd] (\xC,-5.2) -- (\xA,-5.2);
\node[SD/lbl, above] at ({(\xC+\xA)/2},-5.2)
  {5.\ Authorization request (OAuth 2.1 + PKCE)};

%% 6. Auth → Client  (dashed)
\draw[SD/bck] (\xA,-6.2) -- (\xC,-6.2);
\node[SD/lbl, above] at ({(\xC+\xA)/2},-6.2)
  {6.\ Authorization code};

%% 7. Client → Auth  (solid)
\draw[SD/fwd] (\xC,-7.2) -- (\xA,-7.2);
\node[SD/lbl, above] at ({(\xC+\xA)/2},-7.2)
  {7.\ Exchange code + \texttt{code\_verifier} for token};

%% 8. Auth → Client  (dashed)
\draw[SD/bck] (\xA,-8.2) -- (\xC,-8.2);
\node[SD/lbl, above] at ({(\xC+\xA)/2},-8.2)
  {8.\ Access Token};

%% 9. Client → Server  (solid)
\draw[SD/fwd] (\xC,-9.2) -- (\xS,-9.2);
\node[SD/lbl, above] at ({(\xC+\xS)/2},-9.2)
  {9.\ Retry original request with Bearer token};

%% 10. Server → Client  (dashed)
\draw[SD/bck] (\xS,-10.2) -- (\xC,-10.2);
\node[SD/lbl, above] at ({(\xC+\xS)/2},-10.2)
  {10.\ Authorized response (Tool output)};

%% ── Bottom actor boxes (UML convention) ──────────────────────
\node[SD/actor, minimum width=2.3cm] at (\xC,-11.9)
  {MCP Client\\(Agent)};
\node[SD/actor, minimum width=2.3cm] at (\xS,-11.9)
  {MCP Server};
\node[SD/actor, minimum width=2.8cm] at (\xP,-11.9)
  {Protected Resource\\Metadata (PRM)};
\node[SD/actor, minimum width=2.3cm] at (\xA,-11.9)
  {Authorization\\Server};

\end{tikzpicture}
\caption{Standard MCP OAuth 2.1 + PKCE authorization flow (UML sequence
diagram). The MCP client discovers the authorization server via Protected
Resource Metadata, performs the OAuth 2.1 PKCE authorization code flow, and
obtains an access token. Solid arrows are requests; dashed arrows are
responses. Critically, \textbf{this flow covers only the client-to-hub
authorization boundary}. How the hub authorizes its own calls to downstream
worker services is unspecified---the delegation security gap this paper
addresses.}
\label{fig:mcp_std_auth}
\end{figure*}

In multi-service deployments, the AI assistant hub acts as both an MCP server
(receiving tool calls from the agent) and an HTTP client (calling downstream
worker services). This dual role is the hub-and-spoke topology that the MCP
security specification identifies as the confused deputy attack surface.

\subsection{OAuth 2.1 and Bearer Tokens}

OAuth 2.1 \cite{oauth21} is the current revision of the OAuth 2.0 authorization
framework \cite{rfc6749}, mandating PKCE for all public clients and deprecating
the implicit grant. An \textit{access token} is a bearer credential encoding the
authorized subject (\texttt{sub}), audience (\texttt{aud}), and permission set
(\texttt{scope}).

RFC 9068 \cite{rfc9068} profiles access tokens as signed JSON Web Tokens
(JWTs) \cite{rfc7519}, enabling compact, cryptographically verifiable
credential exchange. Four JWT claims are central to this paper.
The \textbf{\texttt{aud}} claim identifies the intended recipient; a receiver
MUST reject tokens not addressed to it \cite{rfc9068}.
The \textbf{\texttt{scope}} claim carries the set of permissions the token grants.
The \textbf{\texttt{jti}} claim provides a unique per-token identifier that
enables revocation and replay prevention.
Finally, the \textbf{\texttt{act}} claim (defined by RFC 8693) records the
identity of the party exercising delegated authority, forming a verifiable
delegation chain.

\subsection{RFC 8693: OAuth 2.0 Token Exchange}

RFC 8693 \cite{rfc8693} defines an HTTP Security Token Service (STS) protocol
for exchanging one token for another with different properties.
We use the \textit{delegation} semantics: the acting party remains
distinguishable from the subject; the new token preserves \texttt{sub} = $u$
while recording the hub in a nested \texttt{act} claim. RFC 8693 also defines
the \textit{scope non-escalation invariant}: a derived token's scope must not
exceed the subject token's scope---the formal basis for P2. The specification
is deployed in production by Microsoft, Auth0, ZITADEL, Authlete, Keycloak,
and Salesforce \cite{rfc8693_deploy}.

\subsection{Token Revocation Stores}

RFC 7009 \cite{rfc7009} defines OAuth 2.0 token revocation. For per-request
one-time-use semantics where a token's \texttt{jti} must be revoked after a
single use---the revocation check must be cross-process and low-latency. The
canonical architecture is an in-memory key-value store with TTL-based expiry:
the record is written after issuance (\textit{register}), queried before each
use (\textit{check}), and overwritten after consumption (\textit{revoke}). Any
implementation satisfying this interface---whether managed (e.g., AWS
ElastiCache, Google Memorystore) or self-hosted (e.g., Redis, DragonflyDB)---
satisfies the requirement. The specific technology choice affects operational
latency and availability, not security semantics.

\subsection{OWASP Generative AI Top 10 (2025)}

The OWASP Top 10 for Large Language Model Applications 2025 \cite{owasp_genai}
identifies the most critical LLM security risks. Two entries are directly
relevant to this work. \textbf{LLM06:2025 Excessive Agency} arises when agents
are granted more permissions or capability than their task requires---this is the
root cause of the scope escalation attack vector studied in this paper.
\textbf{LLM01:2025 Prompt Injection} describes attacker-controlled input that
causes an agent to exercise legitimately authorized tokens in unauthorised ways,
an effect that is significantly amplified when tokens carry excess scopes.


%% ============================================================
\section{Problem Statement}
\label{sec:problem}
%% ============================================================

\subsection{The Token Passthrough Anti-Pattern}

Let $\mathcal{H}$ denote the assistant hub, $\mathcal{W}=\{W_1,\ldots,W_n\}$ the
set of downstream worker services, $u$ the authenticated user, and $\mathcal{A}$
the OAuth authorization server. $\mathcal{A}$ issues a root access token:

\begin{equation}
T_0 = \mathrm{JWT}\!\left(\begin{array}{l}
  \texttt{sub} = u,\quad \texttt{aud} = \mathcal{H},\\[2pt]
  \texttt{scope} = S_0 = \textstyle\bigcup_{i}\, s_i,\quad
  \texttt{jti} = j_0
\end{array}\right)
\label{eq:root}
\end{equation}

where $S_0$ is the union of all scopes required by any worker. In the token
passthrough pattern, $\mathcal{H}$ forwards $T_0$ unchanged to every worker:

\begin{equation}
\mathrm{call}(W_i,\; T_0) \quad \forall\, i \in \{1, \ldots, n\}
\label{eq:passthrough}
\end{equation}

Figure~\ref{fig:passthrough} visualises this topology.

%%  --- FIGURE 2: Token Passthrough Architecture ----
\begin{figure}[t]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[
  every node/.style={font=\small},
  x=1cm, y=1cm,
]
%% Nodes
\node[arch/user]  (alice) at (0, 2.5)   {Alice\\(User)};
\node[arch/oauth] (hub)   at (3.5, 2.5)
  {Hub $\mathcal{H}$\\{\footnotesize aud=$\mathcal{H}$, scope=$S_0$}};
\node[arch/worker](w1)    at (7.5, 4.5)  {$W_1$\\Calendar};
\node[arch/worker](w2)    at (7.5, 2.5)  {$W_2$\\Docs};
\node[arch/worker](w3)    at (7.5, 0.5)  {$W_3$\\Comms};

%% Alice → Hub
\draw[arch/arr] (alice) -- node[above,font=\scriptsize]{$T_0$} (hub);

%% Hub → Workers (same token)
\draw[arch/arr, color=C_WARN]
  (hub) -- node[above,sloped,font=\scriptsize]{$T_0$} (w1);
\draw[arch/arr, color=C_WARN]
  (hub) -- node[above,font=\scriptsize]{$T_0$} (w2);
\draw[arch/arr, color=C_WARN]
  (hub) -- node[below,sloped,font=\scriptsize]{$T_0$} (w3);

%% Attack annotations (below each arrow, right side)
\node[font=\scriptsize, color=C_WARN, text width=2.6cm, align=left]
  at (9.4, 4.5)
  {\textbf{A2}: aud=$\mathcal{H}$,\\accepted at $W_1$};
\node[font=\scriptsize, color=C_WARN, text width=2.6cm, align=left]
  at (9.4, 2.5)
  {\textbf{A1}: all scopes\\visible at $W_2$};
\node[font=\scriptsize, color=C_WARN, text width=2.6cm, align=left]
  at (9.4, 0.5)
  {\textbf{A3}: jti never\\revoked};

%% A4 annotation under hub
\node[font=\scriptsize, color=C_WARN, align=center]
  at (3.5, 0.8)
  {\textbf{A4}: no \texttt{act} claim — 0\% attribution};

%% Border box indicating vulnerability
\draw[draw=C_WARN, dashed, thick, rounded corners=6pt]
  (-0.6, -0.1) rectangle (7.0, 5.3);
\node[font=\scriptsize\bfseries, color=C_WARN] at (3.2, 5.05)
  {Token Passthrough (Insecure)};
\end{tikzpicture}
}
\caption{Token passthrough anti-pattern: the hub forwards $T_0$ (with
hub-bound audience and full scope $S_0$) \emph{unchanged} to every worker,
simultaneously enabling all four attacks (A1--A4) \cite{mcp_security,cosai_mcp}.}
\label{fig:passthrough}
\end{figure}

\subsection{Three Violated Security Properties}

\begin{definition}[\textbf{Audience Binding, P1}]
A token $T$ satisfies audience binding for worker $W_i$ iff
$\mathrm{aud}(T) = \mathrm{AUDIENCE}(W_i)$.
\end{definition}

\textbf{Violation:} $T_0$ carries $\texttt{aud} = \mathcal{H}$. When presented
to $W_i \neq \mathcal{H}$, RFC~9068 requires rejection. In practice, many
implementations omit this check.

\begin{definition}[\textbf{Scope Attenuation, P2}]
A token $T$ satisfies scope attenuation for $W_i$ iff
$\mathrm{scope}(T) \subseteq s_i$, where $s_i$ is the minimum scope $W_i$
requires to execute its function.
\end{definition}

\textbf{Violation:} $T_0$ carries the union of all service scopes,
$S_0 = \bigcup_j s_j$. Each worker therefore receives permissions intended
for every other service, violating the principle of least privilege. A
compromised worker inherits the full authority of the entire system.

\begin{definition}[\textbf{Delegation Chain Visibility, P3}]
A token $T$ satisfies chain visibility iff the \texttt{act} claim of $T$
contains a cryptographically verifiable record of \emph{every} agent that
handled $T$ on behalf of subject $u$ (complete chain, all hops), not merely
the immediate actor.
\end{definition}

\textbf{Violation:} $T_0$ carries no \texttt{act} claim. Every audit record
contains only $\texttt{sub} = u$. Calls originating from different agents are
indistinguishable in the audit trail.

\subsection{The Confused Deputy Problem in MCP}

Hardy's 1988 analysis \cite{hardy1988} applies directly. The \textbf{principal}
is user $u$, who delegated authority $T_0$ to $\mathcal{H}$. The \textbf{deputy}
is hub $\mathcal{H}$, which holds $T_0$ and calls workers on the user's behalf.
The \textbf{confused action} occurs when $\mathcal{H}$ presents $T_0$ to
worker $W_i$, where it carries capabilities that $u$ never intended for $W_i$.
The \textbf{adversary} is any agent or prompt injection payload that exploits
the excess authority embedded in $T_0$.

The MCP specification \cite{mcp_security} states: ``Servers SHOULD NOT pass
tokens received from clients to other servers, as this creates confused deputy
risks.''

\subsection{Four Concrete Attack Vectors}

\begin{table}[h]
\caption{Formal attack vector definitions. Each attack exploits the absence of
one or more security properties (P1~=~audience binding, P2~=~scope attenuation,
P3~=~delegation chain visibility). The ``Requires'' column identifies which
property must be absent for the attack to succeed.}
\label{tab:attack_formal}
\centering
\footnotesize
\setlength{\tabcolsep}{3pt}
\begin{tabularx}{\columnwidth}{l X c}
\toprule
\textbf{ID} & \textbf{Mechanism} & \textbf{Requires} \\
\midrule
A1 & Agent calls a privileged tool using a token without the required scope
     & $\neg$\,P2 \\
A2 & Token obtained for service $W_i$ is replayed at $W_j\;(i\neq j)$
     & $\neg$\,P1 \\
A3 & Captured child token is replayed for a second unauthorised call
     & No revocation \\
A4 & Audit log lacks delegation chain; calls cannot be attributed to agent
     & $\neg$\,P3 \\
\bottomrule
\end{tabularx}
\end{table}


%% ============================================================
\section{The RFC 8693 NAC Architectural Pattern}
\label{sec:solution}
%% ============================================================

%% ---- FIGURE 5: Side-by-side sequence comparison (full width) ----
%% Placed at the very top of the section per IEEEtran float rules.
%% Left panel: token passthrough (insecure baseline).
%% Right panel: RFC 8693 NAC (secure pattern).
\begin{figure*}[!ht]
\centering
\begin{tikzpicture}[
  every node/.style={font=\scriptsize},
  x=1cm, y=1cm,
]

%% ━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━
%%  LEFT PANEL — Token Passthrough (Insecure Baseline)
%% ━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━
\pgfmathsetmacro{\lxU}{1.0}  %% User
\pgfmathsetmacro{\lxH}{3.4}  %% Hub
\pgfmathsetmacro{\lxW}{6.2}  %% Workers (W_1 … W_k combined)

%% Panel title
\node[font=\small\bfseries, color=C_WARN] at (3.65, 0.65)
  {Token Passthrough (Insecure)};

%% Actor boxes
\node[SD/actor, minimum width=1.8cm] at (\lxU,0) {User $u$};
\node[SD/actor, minimum width=1.8cm] at (\lxH,0) {Hub $\mathcal{H}$};
\node[SD/actor, minimum width=2.1cm] at (\lxW,0)
  {Workers\\$W_1 \ldots W_k$};

%% Lifelines
\draw[SD/life] (\lxU,-0.30) -- (\lxU,-8.2);
\draw[SD/life] (\lxH,-0.30) -- (\lxH,-8.2);
\draw[SD/life] (\lxW,-0.30) -- (\lxW,-8.2);

%% 1. User → Hub: root token T0
\draw[SD/fwd] (\lxU,-1.1) -- (\lxH,-1.1);
\node[SD/lbl, above] at ({(\lxU+\lxH)/2},-1.1)
  {Bearer $T_0$\;(\texttt{aud}$=\mathcal{H}$,\;scope$=S_0$)};

%% 2. Hub → Workers: same T0 forwarded unchanged (first worker)
\draw[SD/fwd, color=C_WARN, thick] (\lxH,-2.4) -- (\lxW,-2.4);
\node[SD/lbl, above, color=C_WARN] at ({(\lxH+\lxW)/2},-2.4)
  {$T_0$ unchanged \textemdash{} all workers};
\node[font=\tiny, color=C_WARN, align=center]
  at ({(\lxH+\lxW)/2},-2.85)
  {\texttt{aud}$=\mathcal{H}$,\;scope$=S_0$ (full)};

%% Vertical dots (represents repeated passthrough to each worker)
\node[font=\large, color=gray] at ({(\lxH+\lxW)/2},-3.7) {$\vdots$};

%% 3. Hub → Workers: same T0 (k-th call)
\draw[SD/fwd, color=C_WARN, thick] (\lxH,-4.6) -- (\lxW,-4.6);
\node[SD/lbl, above, color=C_WARN] at ({(\lxH+\lxW)/2},-4.6)
  {$T_0$ unchanged \textemdash{} $k$-th worker};

%% Attack consequence box
\node[draw=C_WARN, fill=C_WARN!8, rounded corners=3pt,
      font=\tiny, align=left, inner sep=4pt, text width=5.6cm]
  at (3.65,-6.5)
  {\textbf{Consequences:}\\
   A1: any worker sees full scope $S_0$\\
   A2: $T_0$ accepted at any $W_j$ (wrong audience)\\
   A3: $T_0$ reusable across sessions\\
   A4: no delegation chain; calls unattributable};

%% Left panel border
\draw[draw=C_WARN!70, dashed, thick, rounded corners=5pt]
  (0.0,-8.2) rectangle (7.3,0.35);

%% ━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━
%%  VERTICAL DIVIDER
%% ━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━
\draw[thick, gray!45] (7.65, 0.8) -- (7.65,-8.5);

%% ━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━
%%  RIGHT PANEL — RFC 8693 NAC (Secure Pattern)
%% ━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━━
\pgfmathsetmacro{\rxU}{8.8}   %% User
\pgfmathsetmacro{\rxH}{11.0}  %% Hub
\pgfmathsetmacro{\rxA}{13.7}  %% Auth Server
\pgfmathsetmacro{\rxW}{16.4}  %% Workers

%% Panel title
\node[font=\small\bfseries, color=C_PAR] at (12.7, 0.65)
  {RFC 8693 NAC (Secure)};

%% Actor boxes
\node[SD/actor, minimum width=1.8cm] at (\rxU,0) {User $u$};
\node[SD/actor, minimum width=1.8cm] at (\rxH,0) {Hub $\mathcal{H}$};
\node[SD/actor, minimum width=2.4cm] at (\rxA,0)
  {Auth Server\\(RFC 8693)};
\node[SD/actor, minimum width=2.1cm] at (\rxW,0)
  {Workers\\$W_1 \ldots W_k$};

%% Lifelines
\draw[SD/life] (\rxU,-0.30) -- (\rxU,-8.2);
\draw[SD/life] (\rxH,-0.30) -- (\rxH,-8.2);
\draw[SD/life] (\rxA,-0.30) -- (\rxA,-8.2);
\draw[SD/life] (\rxW,-0.30) -- (\rxW,-8.2);

%% 1. User → Hub: root token T0
\draw[SD/fwd] (\rxU,-1.1) -- (\rxH,-1.1);
\node[SD/lbl, above] at ({(\rxU+\rxH)/2},-1.1)
  {Bearer $T_0$\;(\texttt{aud}$=\mathcal{H}$,\;scope$=S_0$)};

%% 2. Concurrent exchange — "par" combined fragment
\draw[SD/par] (\rxH-0.25,-1.7) rectangle (\rxA+0.35,-4.1);
\node[font=\tiny\bfseries, fill=white, inner sep=1.5pt, anchor=north west]
  at (\rxH-0.25,-1.7) {par\;($i = 1\ldots k$, concurrent)};

%% Hub → Auth: exchange request (per worker)
\draw[SD/fwd, color=C_PAR!80!black] (\rxH,-2.35) -- (\rxA,-2.35);
\node[SD/lbl, above, color=C_PAR!80!black] at ({(\rxH+\rxA)/2},-2.35)
  {\texttt{POST /token/exchange}\;(\texttt{aud}$=W_i$,\;scope$=s_i$)};

\node[font=\large, color=gray] at ({(\rxH+\rxA)/2},-3.0) {$\vdots$};
\node[font=\tiny, color=gray, align=center]
  at ({(\rxH+\rxA)/2},-3.45) {one request per worker, fired simultaneously};

%% Auth Server: validates, signs child tokens
\node[draw=C_BASE!55, fill=C_BASE!10, rounded corners=2pt,
      font=\tiny, align=center, inner sep=3pt]
  at (\rxA,-3.85)
  {enforce $s_i \!\subseteq\! S_0$; sign $T_i$; register jti$_i$};

%% Auth → Hub: child tokens returned
\draw[SD/bck] (\rxA,-4.55) -- (\rxH,-4.55);
\node[SD/lbl, above] at ({(\rxH+\rxA)/2},-4.55)
  {$T_1,\ldots,T_k$: \texttt{aud}$=W_i$,\;scope$=s_i$,\;\texttt{act}$=\{\text{sub}:\mathcal{H}\}$};

%% 3. Worker invocation — "loop" combined fragment
\draw[SD/par] (\rxH-0.25,-5.1) rectangle (\rxW+0.25,-7.7);
\node[font=\tiny\bfseries, fill=white, inner sep=1.5pt, anchor=north west]
  at (\rxH-0.25,-5.1) {loop\;($i = 1\ldots k$)};

%% Hub → Workers: Ti (unique per worker)
\draw[SD/fwd, color=C_PAR] (\rxH,-5.75) -- (\rxW,-5.75);
\node[SD/lbl, above, color=C_PAR] at ({(\rxH+\rxW)/2},-5.75)
  {Bearer $T_i$\;(\texttt{aud}$=W_i$,\;scope$=s_i$,\;\texttt{act} chain)};

%% Worker validation annotation
\node[draw=C_PAR!55, fill=C_PAR!10, rounded corners=2pt,
      font=\tiny, align=left, inner sep=3pt]
  at (\rxW,-6.55)
  {validate: \texttt{aud}, scope, act,\\jti (revoke after use)};

%% Workers → Hub: tool result
\draw[SD/bck] (\rxW,-7.3) -- (\rxH,-7.3);
\node[SD/lbl, above] at ({(\rxH+\rxW)/2},-7.3) {tool result};

%% Right panel border
\draw[draw=C_PAR!70, dashed, thick, rounded corners=5pt]
  (7.8,-8.2) rectangle (17.5,0.35);

\end{tikzpicture}
\caption{Side-by-side sequence comparison of token passthrough (left) and the
RFC~8693 NAC pattern (right). In the baseline flow, the hub forwards the
unchanged root token $T_0$ to all workers, simultaneously exposing attacks
A1--A4. In the NAC flow, the hub exchanges $T_0$ for $k$ per-service child
tokens at the authorization server (concurrent, \textit{par} fragment) before
invoking any worker; each child token carries a unique audience, an attenuated
scope, a verifiable \texttt{act} chain, and a one-time-use \texttt{jti},
blocking all four attacks by construction.}
\label{fig:seq_comparison}
\end{figure*}

\subsection{Pattern Overview}

Figure~\ref{fig:seq_comparison} contrasts the two flows at a glance.
Intuitively, the NAC pattern replaces a single all-access key with a set of
single-use, room-specific keys: instead of handing the same master key to every
service, the hub asks a trusted authority to cut a new key for each service,
valid only for that service, carrying only the permissions it needs, and expiring
after one use. The authority records in the key itself a chain of custody showing
which agent made the request on behalf of which user.

More precisely: Before invoking worker $W_i$, the orchestrating hub presents its
current token to the authorization server's token exchange endpoint and obtains a
fresh, narrowly scoped child token:

\begin{equation}
T_i = \mathrm{Exchange}\!\left(
  T_{\mathrm{parent}},\;
  \mathrm{AUDIENCE}(W_i),\;
  s_i,\;
  \mathcal{H}
\right)
\label{eq:exchange}
\end{equation}

where $T_i$ is constructed by the authorization server as:

\begin{equation}
T_i = \mathrm{JWT}\!\left(\begin{array}{l}
  \texttt{sub} = u, \\[2pt]
  \texttt{aud} = \mathrm{AUDIENCE}(W_i), \\[2pt]
  \texttt{scope} = s_i \;\subseteq\; \mathrm{scope}(T_{\mathrm{parent}}), \\[2pt]
  \texttt{jti} = j_i \;\text{(fresh, unique)}, \\[2pt]
  \texttt{act} = \{\,\texttt{sub}\text{:}\,\mathcal{H},\;
                    \texttt{act}\text{:}\,T_{\mathrm{parent}}.\texttt{act}\,\}
\end{array}\right)
\label{eq:child}
\end{equation}

This construction directly restores all three violated properties: the
\texttt{aud} satisfies P1, the scope satisfies P2, and the \texttt{act} chain
satisfies P3. Figure~\ref{fig:nac_arch} shows the NAC architecture, and
Figure~\ref{fig:nac_seq} illustrates the complete token exchange sequence.

Listing~\ref{lst:child} shows the reference implementation's child token
payload construction, which maps directly to Equation~(\ref{eq:child}).

\begin{lstlisting}[language=Python, label=lst:child,
  caption={Child token construction (\texttt{nac\_common.py}).
  Each field enforces one security property.}]
payload = {
  "sub":   parent["sub"],           # subject preserved
  "aud":   new_audience,            # P1: per-worker audience
  "scope": scope_to_str(new_scope), # P2: attenuated scope
  "jti":   str(uuid.uuid4()),       # one-time-use ID
  "act":   {"sub": actor,           # P3: delegation chain
             "act": parent.get("act")},
}
child_token = jwt.encode(payload, signing_key, algorithm="RS256")
\end{lstlisting}

%% ---- FIGURE 3: NAC Architecture (single column, fixed) ----
\begin{figure}[t]
\centering
\resizebox{\columnwidth}{!}{%
\begin{tikzpicture}[
  every node/.style={font=\small},
  x=1cm, y=1cm,
]
%% Nodes
\node[arch/user]  (alice)  at (0, 3.0)
  {Alice};
\node[arch/oauth] (auth)   at (3.5, 5.5)
  {Auth Server\\{\footnotesize(RFC 8693)}};
\node[arch/hub]   (hub)    at (3.5, 3.0)
  {Hub $\mathcal{H}$};
\node[arch/worker](w1)     at (7.5, 5.2)
  {$W_1$\\Calendar};
\node[arch/worker](w2)     at (7.5, 3.2)
  {$W_2$\\Docs};
\node[arch/worker](w3)     at (7.5, 1.2)
  {$W_3$\\Comms};

%% Alice → Hub
\draw[arch/arr] (alice) -- node[above,font=\scriptsize]{$T_0$} (hub);

%% Hub ↔ Auth Server (exchange)
\draw[arch/arr, color=C_SEC, bend left=18]
  (hub.north) to node[right,font=\scriptsize,pos=0.5]
    {\texttt{/exchange} $\times k$} (auth.south);
\draw[arch/darr, color=C_BASE, bend left=18]
  (auth.south) to node[left,font=\scriptsize,pos=0.5]
    {$T_1, T_2, T_3$} (hub.north);

%% Hub → Workers (different tokens, colour-coded)
\draw[arch/arr, color=C_PAR]
  (hub) -- node[above,sloped,font=\scriptsize]{$T_1$} (w1);
\draw[arch/arr, color=C_PAR]
  (hub) -- node[above,font=\scriptsize]{$T_2$} (w2);
\draw[arch/arr, color=C_PAR]
  (hub) -- node[below,sloped,font=\scriptsize]{$T_3$} (w3);

%% Token property labels next to workers
\node[font=\tiny, color=C_PAR, text width=2.4cm, align=left]
  at (9.4, 5.6)
  {aud=$W_1$, scope=$s_1$\\act=\{sub:$\mathcal{H}$\}};
\node[font=\tiny, color=C_PAR, text width=2.4cm, align=left]
  at (9.4, 3.2)
  {aud=$W_2$, scope=$s_2$\\act=\{sub:$\mathcal{H}$\}};
\node[font=\tiny, color=C_PAR, text width=2.4cm, align=left]
  at (9.4, 1.2)
  {aud=$W_3$, scope=$s_3$\\act=\{sub:$\mathcal{H}$\}};

%% Green border indicating security
\draw[draw=C_PAR, dashed, thick, rounded corners=6pt]
  (-0.6, -0.1) rectangle (7.0, 6.2);
\node[font=\scriptsize\bfseries, color=C_PAR!80!black]
  at (3.2, 6.0) {NAC-Secured (RFC 8693)};
\end{tikzpicture}
}
\caption{RFC 8693 NAC pattern: the hub exchanges the root token $T_0$ for
per-service child tokens at the authorization server before each worker call.
Each $T_i$ carries a unique \texttt{aud}, attenuated \texttt{scope}, and
nested \texttt{act} chain, satisfying P1--P3 simultaneously.}
\label{fig:nac_arch}
\end{figure}

\subsection{Concurrent Issuance for Multiple Workers}

Because child tokens for $W_1, W_2, \ldots, W_k$ are issued from the same parent
token and are mutually independent (different audience, different scope, no
ordering constraint), they can be issued concurrently:

\begin{multline}
(T_1, T_2, \ldots, T_k) = \\
\mathrm{ConcurrentExchange}\!\left(
  T_{\mathrm{root}},\;
  \bigl\{\,(\mathrm{aud}_i,\; s_i)\,\bigr\}_{i=1}^{k}
\right)
\label{eq:concurrent}
\end{multline}

Concurrent issuance reduces wall-clock time from $k \times t_{\mathrm{exchange}}$
to $\approx t_{\mathrm{exchange}}$ (the maximum across concurrent requests),
subject to authorization server capacity. This is the basis for the
``parallel estimate'' in Section~\ref{sec:eval}.

\begin{proposition}[\textbf{Overhead Bound}]
\label{prop:overhead}
For $k$ independent workers, sequential NAC overhead is $O(k)$ with a
constant equal to one RSA signing operation plus one revocation store write
per hop. Under concurrent issuance, wall-clock overhead collapses to $O(1)$
relative to $k$, bounded by $t_{\mathrm{sign}} + t_{\mathrm{store}}$ --- the
cost of a single exchange --- regardless of how many workers are invoked.
\end{proposition}
\begin{proof}
Each child token exchange is a function solely of
$(T_{\mathrm{root}},\, \mathrm{aud}_i,\, s_i)$. Because no exchange reads or
modifies the output of any other exchange, the $k$ operations are fully
independent --- there is no cross-exchange data dependency. Sequential
execution therefore costs exactly $k \cdot (t_{\mathrm{sign}} +
t_{\mathrm{store}} + t_{\mathrm{http}})$; concurrent execution via
\texttt{asyncio.gather} or equivalent reduces the wall-clock component to
$\max_i(t_{\mathrm{exchange},i})$, which approaches $t_{\mathrm{sign}} +
t_{\mathrm{store}}$ when exchanges are dispatched simultaneously. The signing
and store operations are cryptographically fixed-cost (RSA-2048 sign,
sub-millisecond Redis write); only $t_{\mathrm{http}}$ varies with deployment
topology and is eliminable via co-location of the authorization server with
the hub.
\end{proof}

Listing~\ref{lst:gather} shows the hub implementation that realises
Equation~(\ref{eq:concurrent}) — all four RFC~8693 exchanges are dispatched
simultaneously via \texttt{asyncio.gather}, collapsing sequential latency
$k \times t_{\mathrm{exchange}}$ to a single round-trip.

\begin{lstlisting}[language=Python, label=lst:gather,
  caption={Concurrent RFC 8693 exchange (\texttt{assistant\_server.py});
  all four tokens are issued in parallel, achieving $O(1)$ wall-clock overhead.}]
cal, docs, email, slack = await asyncio.gather(
    _get_worker_token(root, "calendar", ["calendar:read"]),
    _get_worker_token(root, "docs",     ["docs:read"]),
    _get_worker_token(root, "comms",    ["email:send"]),
    _get_worker_token(root, "comms",    ["slack:write"]),
)
\end{lstlisting}

\subsection{N-Hop Delegation Chains}

For hierarchical delegation---where $W_i$ must itself call a further downstream
service---the pattern applies recursively:

\begin{equation}
T_k = \mathrm{Exchange}(T_{k-1},\;
  \mathrm{AUDIENCE}(W_k),\; s_k,\; W_{k-1})
\label{eq:nhop}
\end{equation}

The \texttt{act} claim grows by one level per hop, recording the complete
delegation history; the IETF Cross-Domain Attributes in Authentication and
Authorization Mechanisms (CAAM) draft \cite{ietf_caam} formalises this
multi-hop trust model. A depth limit prevents unbounded chain growth or
circular references. We demonstrate a 3-hop chain:
$\text{alice} \to \mathcal{H} \to W_{\mathrm{cal}} \to W_{\mathrm{ext}}$.

\subsection{NAC Token Exchange Sequence}

Figure~\ref{fig:nac_seq} provides the full UML sequence diagram for the NAC
token exchange flow. The diagram uses four actors---Alice, the hub
$\mathcal{H}$, the authorization server, and a representative worker
$W_i$---to show the general $k$-worker pattern. Phase~2 uses a ``par''
combined fragment for the concurrent exchanges; Phase~3 uses a ``loop''
fragment for the sequential worker invocations, each with its own
validation and \texttt{jti} revocation step.

%% ---- FIGURE 4: NAC Exchange Sequence Diagram (full width) ----
%% Redesigned: 4 actors (Alice, Hub, Auth Server, Worker Wi) with
%% proper spacing matching the reference MCP OAuth flow style.
%% Uses par fragment for concurrent exchanges, loop fragment for
%% worker calls, and activation boxes for worker validation steps.
\begin{figure*}[!ht]
\centering
\begin{tikzpicture}[
  every node/.style={font=\scriptsize},
  x=1cm, y=1cm,
]

%% ── Actor x-positions ──────────────────────────────────────────────────────
%% 4 actors across ~17cm textwidth — matches the reference image spacing style
%% Each actor is ~2.6cm wide; gaps of ~4cm give clear, uncrossed lifelines.
\pgfmathsetmacro{\xAl}{1.5}   %% Alice (User)
\pgfmathsetmacro{\xHu}{6.0}   %% Hub H
\pgfmathsetmacro{\xAu}{11.5}  %% Auth Server (RFC 8693)
\pgfmathsetmacro{\xWk}{16.5}  %% Worker W_i  (represents all k workers)

%% ── Top actor boxes ────────────────────────────────────────────────────────
\node[SD/actor, minimum width=2.4cm] (A1) at (\xAl, 0)
  {Alice\\(User)};
\node[SD/actor, minimum width=2.4cm] (H1) at (\xHu, 0)
  {Hub $\mathcal{H}$};
\node[SD/actor, minimum width=3.0cm] (Au1) at (\xAu, 0)
  {Auth Server\\(RFC 8693)};
\node[SD/actor, minimum width=3.0cm] (W1) at (\xWk, 0)
  {Worker $W_i$\\{\tiny$(i = 1 \ldots k)$}};

%% ── Lifelines ──────────────────────────────────────────────────────────────
\draw[SD/life] (\xAl,-0.32) -- (\xAl,-17.0);
\draw[SD/life] (\xHu,-0.32) -- (\xHu,-17.0);
\draw[SD/life] (\xAu,-0.32) -- (\xAu,-17.0);
\draw[SD/life] (\xWk,-0.32) -- (\xWk,-17.0);

%% ══════════════════════════════════════════════════════════════════════════
%% PHASE 1 — Root token delivery
%% ══════════════════════════════════════════════════════════════════════════
\draw[SD/fwd] (\xAl,-1.1) -- (\xHu,-1.1);
\node[SD/lbl, above] at ({(\xAl+\xHu)/2},-1.1)
  {1.\ Bearer $T_0$ in \texttt{Authorization} header
   \quad (\texttt{aud}$=\mathcal{H}$,\ \texttt{scope}$=S_0$)};

%% Phase label
\node[font=\tiny\itshape, color=gray] at ({(\xAl+\xHu)/2},-1.6)
  {Phase 1: Root token};

%% ══════════════════════════════════════════════════════════════════════════
%% PHASE 2 — Concurrent RFC 8693 exchanges
%% ══════════════════════════════════════════════════════════════════════════
\node[font=\tiny\itshape, color=gray] at ({(\xHu+\xAu)/2},-2.1)
  {Phase 2: Concurrent token exchange};

%% "par" combined-fragment box (dashed rectangle with label)
\draw[SD/par] (\xHu-0.4,-2.4) rectangle (\xAu+0.6,-7.2);
\node[font=\tiny\bfseries, fill=white, inner sep=1.5pt, anchor=north west]
  at (\xHu-0.4,-2.4) {par\ (concurrent,\ $i=1\ldots k$)};

%% 2a — first exchange
\draw[SD/fwd, color=C_PAR!80!black] (\xHu,-3.0) -- (\xAu,-3.0);
\node[SD/lbl, above] at ({(\xHu+\xAu)/2},-3.0)
  {2a.\ \texttt{POST /token/exchange}
   (\texttt{aud}=$W_1$,\ \texttt{scope}=$s_1$,\ actor=$\mathcal{H}$)};

%% vertical dots (represents 2b … 2(k-1))
\node[font=\small, color=gray] at ({(\xHu+\xAu)/2},-4.0) {$\vdots$};
\node[SD/lbl] at ({(\xHu+\xAu)/2},-4.4) {\textit{(one request per worker, fired simultaneously)}};

%% 2k — last exchange
\draw[SD/fwd, color=C_PAR!80!black] (\xHu,-5.2) -- (\xAu,-5.2);
\node[SD/lbl, above] at ({(\xHu+\xAu)/2},-5.2)
  {2k.\ \texttt{POST /token/exchange}
   (\texttt{aud}=$W_k$,\ \texttt{scope}=$s_k$,\ actor=$\mathcal{H}$)};

%% Auth server processing note
\node[SD/lbl, fill=C_BASE!12, draw=C_BASE!45, rounded corners=3pt,
      font=\tiny, align=center, minimum width=3.4cm]
  at (\xAu,-6.4)
  {enforce $s_i \subseteq \mathrm{scope}(T_0)$\\
   sign $T_i$;\ register jti$_i$};

%% 3 — child tokens returned
\draw[SD/bck] (\xAu,-8.0) -- (\xHu,-8.0);
\node[SD/lbl, above] at ({(\xHu+\xAu)/2},-8.0)
  {3.\ $T_1, T_2, \ldots, T_k$\;—\;each: unique \texttt{aud},
   attenuated \texttt{scope}, \texttt{act}=\{\texttt{sub}:$\mathcal{H}$\}};

%% ══════════════════════════════════════════════════════════════════════════
%% PHASE 3 — Worker invocation (loop)
%% ══════════════════════════════════════════════════════════════════════════
\node[font=\tiny\itshape, color=gray] at ({(\xHu+\xWk)/2},-8.6)
  {Phase 3: Worker invocation};

%% "loop" combined-fragment box
\draw[SD/par] (\xHu-0.4,-8.9) rectangle (\xWk+0.4,-14.8);
\node[font=\tiny\bfseries, fill=white, inner sep=1.5pt, anchor=north west]
  at (\xHu-0.4,-8.9) {loop\ ($i = 1 \ldots k$)};

%% 4 — Hub sends Ti to Wi
\draw[SD/fwd] (\xHu,-9.6) -- (\xWk,-9.6);
\node[SD/lbl, above] at ({(\xHu+\xWk)/2},-9.6)
  {4.\ Bearer $T_i$
   (\texttt{aud}=$W_i$,\ \texttt{scope}=$s_i$)};

%% Worker validation box (activation on lifeline)
\node[SD/lbl, fill=C_PAR!15, draw=C_PAR!55, rounded corners=3pt,
      font=\tiny, align=left, minimum width=3.6cm, inner sep=4pt]
  at (\xWk,-11.3)
  {{\bfseries Validate:}\\
   \texttt{aud}($T_i$) $= W_i$ \quad(P1)\\
   \texttt{scope}($T_i$) $\subseteq s_i$ \quad(P2)\\
   \texttt{act}\ chain trusted \quad(P3)\\
   jti$_i \notin$ revocation store};

%% 5 — tool result
\draw[SD/bck] (\xWk,-12.3) -- (\xHu,-12.3);
\node[SD/lbl, above] at ({(\xHu+\xWk)/2},-12.3)
  {5.\ Tool result};

%% Worker post-use revocation box
\node[SD/lbl, fill=C_WARN!12, draw=C_WARN!40, rounded corners=3pt,
      font=\tiny, align=center, minimum width=3.2cm]
  at (\xWk,-13.4)
  {revoke jti$_i$ in store\\
   (one-time-use enforced)};

%% revocation written
\draw[SD/bck, color=C_WARN!55, dashed] (\xWk,-14.3) -- (\xHu,-14.3);
\node[SD/lbl, above] at ({(\xHu+\xWk)/2},-14.3)
  {(jti$_i$ marked revoked)};

%% 6 — Hub aggregates and responds to Alice
\draw[SD/bck] (\xHu,-15.5) -- (\xAl,-15.5);
\node[SD/lbl, above] at ({(\xAl+\xHu)/2},-15.5)
  {6.\ Aggregated response to Alice};

%% ── Bottom actor boxes (UML convention) ────────────────────────────────────
\node[SD/actor, minimum width=2.4cm] at (\xAl,-16.6) {Alice\\(User)};
\node[SD/actor, minimum width=2.4cm] at (\xHu,-16.6) {Hub $\mathcal{H}$};
\node[SD/actor, minimum width=3.0cm] at (\xAu,-16.6) {Auth Server\\(RFC 8693)};
\node[SD/actor, minimum width=3.0cm] at (\xWk,-16.6)
  {Worker $W_i$\\{\tiny$(i = 1 \ldots k)$}};

\end{tikzpicture}
\caption{RFC 8693 NAC token exchange sequence (UML, full pattern).
\textbf{Phase~1}: Alice supplies root token $T_0$ to the hub.
\textbf{Phase~2}: The hub fires $k$ concurrent RFC~8693 exchange requests
(``par'' combined fragment); the authorization server enforces scope
attenuation ($s_i \subseteq \mathrm{scope}(T_0)$) and returns per-service
child tokens, each with a unique audience, attenuated scope, and a nested
\texttt{act} claim recording $\mathcal{H}$ as the acting party.
\textbf{Phase~3}: Each worker ($W_i$) receives only its own token,
validates all four security properties (P1--P3 plus JTI freshness), and
revokes the \texttt{jti} after first use, preventing replay.
Solid arrows are requests; dashed arrows are responses.
$W_i$ represents each worker in sequence; the diagram generalises to
$k$ workers.}
\label{fig:nac_seq}
\end{figure*}

\subsection{Security Property Satisfaction}

\begin{theorem}[\textbf{P1 satisfied}]
Worker $W_i$ receives only tokens with $\texttt{aud}=\mathrm{AUDIENCE}(W_i)$.
Any replayed token from $W_j\;(j\neq i)$ fails the audience check.
\end{theorem}
\begin{proof}
By construction (Equation~\ref{eq:child}), the authorization server sets
$\texttt{aud}=\mathrm{AUDIENCE}(W_i)$. Workers hold only the public signing
key and cannot produce or modify tokens. Any token with
$\texttt{aud}\neq\mathrm{AUDIENCE}(W_i)$ fails JWT audience validation.
\end{proof}

\begin{theorem}[\textbf{P2 satisfied}]
The authorization server enforces $s_i \subseteq \mathrm{scope}(T_{\mathrm{parent}})$
at exchange time. Any request for scope $s^* \not\subseteq
\mathrm{scope}(T_{\mathrm{parent}})$ is rejected before any child token is issued.
\end{theorem}
\begin{proof}
The exchange endpoint computes $\delta = s^* \setminus
\mathrm{scope}(T_{\mathrm{parent}})$. If $\delta\neq\emptyset$, it returns HTTP
400. No child token is signed; the attacker gains no elevated capability.
This follows directly from the RFC 8693 scope non-escalation invariant
\cite{rfc8693}.
\end{proof}

\begin{theorem}[\textbf{P3 satisfied}]
The \texttt{act} claim in each child token is a cryptographically signed record
of every delegation hop, enabling full attribution.
\end{theorem}
\begin{proof}
The \texttt{act} claim is embedded by the authorization server at issuance and
is part of the RS256-signed payload. Any alteration invalidates the signature.
Verifying parties walk the nested \texttt{act} chain and confirm each actor
against a set of trusted identities.
\end{proof}

\begin{theorem}[\textbf{Token Replay Prevented}]
Under the NAC pattern with per-request \texttt{jti} revocation, token $T_i$
with identifier $j_i$ is rejected on any second presentation.
\end{theorem}
\begin{proof}
After first successful validation, $W_i$ writes $j_i$ to the revocation store
with status \textit{revoked} (cf.\ RFC~7009 token revocation semantics \cite{rfc7009}). On second presentation, the revocation check
returns \textit{revoked} and $W_i$ returns an error before executing any
business logic. The TTL of the revocation record must exceed the token's
remaining lifetime to prevent races.
\end{proof}

\subsection{Architectural Requirements for Adopters}

The NAC pattern places five requirements on the deployment. The
\textbf{authorization server} must be any OAuth 2.1-compliant server that
exposes a token exchange endpoint conforming to RFC 8693; existing enterprise
identity providers (Keycloak, Auth0, ZITADEL, Authlete, and Microsoft Entra via
the OBO flow) support this natively \cite{rfc8693_deploy}. The
\textbf{orchestrator} must perform one RFC 8693 exchange per worker invocation,
firing exchanges concurrently for independent workers
(Equation~\ref{eq:concurrent}). Each \textbf{worker} must validate the
\texttt{aud}, \texttt{scope}, \texttt{act} chain, and \texttt{jti} on every
incoming request, and must write the \texttt{jti} to the revocation store after
successful use. The \textbf{revocation store} must support sub-millisecond
read/write of \texttt{jti} records shared across processes with TTL-based
expiry. Finally, \textbf{key separation} must be enforced: only the
authorization server holds the RSA private signing key; workers hold the
corresponding public key and are physically prevented from loading the private
key.

No requirement is placed on programming language, web framework, or specific
runtime. The pattern applies equally to Java, C\#, Python, Node.js, Go, or any
language with an HTTP client and JWT library.


%% ============================================================
\section{Reference Implementation}
\label{sec:impl}
%% ============================================================

Theorems~1--4 and Proposition~\ref{prop:overhead} establish the security and
overhead properties of the NAC pattern by construction. To provide empirical
confirmation, we constructed a minimal reference implementation: two parallel
server stacks sharing identical six-component topology, differing only in
whether RFC 8693 exchange and token validation are enforced. The implementation
is intentionally minimal---each component is a single Python file using FastAPI
and standard libraries---so that every security-relevant decision is
immediately auditable and the distance between theory and code is as small as
possible.

\subsection{Component Architecture}

\textbf{Authorization server}: Implements RFC 8693 \texttt{/token/exchange}.
In the secure stack, it validates the subject token, enforces scope attenuation,
issues a signed child token, and registers the new \texttt{jti}. In the baseline,
the endpoint echoes the subject token unchanged.

\textbf{Assistant hub}: Acts as both an MCP server (to the agent) and an HTTP
client (to workers and the authorization server). In the secure stack, it fires
concurrent RFC 8693 exchanges for all workers before any worker call using a
shared HTTP connection pool to amortise TCP overhead across rounds.

\textbf{Worker services}: Four services (Calendar, Documents, Communications,
External-API) each implementing token validation. In the secure stack, all four
properties (P1--P3 plus replay prevention) are enforced and the \texttt{jti} is
revoked after first use. In the baseline, validation is omitted.

\textbf{Revocation store}: We use Redis \cite{redis} as the revocation store
in our reference implementation---a widely deployed in-memory key-value server
that provides the sub-millisecond operation latency required for per-request
validation. Key schema:

\begin{equation}
\texttt{nac:jti:}\langle\text{uuid}\rangle \to
\begin{cases}
  \texttt{"active"}  & \text{TTL} = \text{token lifetime} + 60\,\text{s} \\
  \texttt{"revoked"} & \text{TTL} = 300\,\text{s}
\end{cases}
\end{equation}

Listing~\ref{lst:jti} shows the three JTI operations that implement this
schema. Each operation completes in $\approx 0.1$\,ms (co-located Redis,
binary protocol), contributing only $\approx 2\%$ of total overhead.

\begin{lstlisting}[language=Python, label=lst:jti,
  caption={Redis JTI store: register on issue, revoke on exchange,
  check on each worker request (\texttt{nac\_common.py}, $\approx$0.1\,ms/op).}]
def register_jti(jti, exp):          # called by auth server on issue
    ttl = max(1, int(exp - time.time()) + 60)
    redis.set(f"nac:jti:{jti}", "active", ex=ttl)

def revoke_jti(jti):                 # called after token is exchanged
    redis.set(f"nac:jti:{jti}", "revoked", ex=300)

def is_jti_revoked(jti):             # called by worker before accepting
    return redis.get(f"nac:jti:{jti}") == "revoked"
\end{lstlisting}

Any store satisfying the interface in Section~\ref{sec:solution} may be
substituted; the security semantics are unaffected.

\textbf{Agent planner}: The evaluation uses a rule-based planner for
reproducibility. An optional real LLM planner demonstrates that the NAC pattern
is transparent to the model layer---the LLM receives only tool names and results;
token exchange occurs at the transport layer.

\subsection{Key Security Invariants}

Three invariants are enforced by construction in the reference implementation.
\textbf{Key separation}: only the authorization server process holds the
RSA-2048 private signing key; worker processes reject any attempt to load the
private key, preventing token forgery even under worker compromise.
\textbf{Short-lived child tokens}: child tokens carry a significantly shorter
TTL than the root token, minimising the replay window in the event that a
\texttt{jti} revocation record is lost.
\textbf{Typed error codes}: validation failures return structured string
constants (\texttt{WRONG\_AUDIENCE}, \texttt{SCOPE\_INSUFFICIENT},
\texttt{TOKEN\_REPLAY}), enabling the evaluation harness to distinguish
security rejections from implementation bugs.

Listing~\ref{lst:validate} shows the \texttt{validate\_token} function shared
by all four worker services. Each enforcement flag corresponds to one theorem:
audience binding (Theorem~1), scope attenuation (Theorem~2), and JTI replay
prevention (Theorem~4).

\begin{lstlisting}[language=Python, label=lst:validate,
  caption={Worker token validation (\texttt{nac\_common.py}); each check
  corresponds to one theorem.}]
def validate_token(token, expected_audience, required_scopes,
                   enforce_jti=False):
    # Theorem 1 (P1): raises InvalidAudienceError if aud mismatches
    claims = jwt.decode(token, pub_key, algorithms=["RS256"],
                        audience=expected_audience)
    if enforce_jti and is_jti_revoked(claims["jti"]):
        raise ValueError("TOKEN_REPLAY")        # Theorem 4
    if any(s not in claims["scope"] for s in required_scopes):
        raise ValueError("SCOPE_INSUFFICIENT")  # Theorem 2 (P2)
    return claims
\end{lstlisting}

\subsection{Port Layout}

\begin{table}[h]
\caption{Reference implementation port layout (two isolated stacks
run concurrently; full configuration in the supplementary repository)}
\label{tab:ports}
\centering
\footnotesize
\begin{tabular}{lcccccc}
\toprule
\textbf{Stack} & \textbf{Auth} & \textbf{Hub} & \textbf{Cal}
               & \textbf{Docs} & \textbf{Comms} & \textbf{Ext} \\
\midrule
Baseline (insecure) & 9200 & 9201 & 9202 & 9203 & 9204 & 9205 \\
Secure (NAC)        & 9300 & 9301 & 9302 & 9303 & 9304 & 9305 \\
\bottomrule
\end{tabular}
\end{table}


%% ============================================================
\section{Evaluation}
\label{sec:eval}
%% ============================================================

%% ---- FIGURE: Combined evaluation overview (full width) ----
\begin{figure*}[!t]
\centering
\begin{tikzpicture}
\begin{groupplot}[
  group style={
    group size=2 by 1,
    horizontal sep=1.8cm,
  },
  width=0.46\textwidth,
  height=5.0cm,
]

%% ── Left panel: Attack block rates ──────────────────────────
\nextgroupplot[
  ybar, bar width=10pt,
  symbolic x coords={A1,A2,A3,A4},
  xtick=data,
  ymin=0, ymax=115,
  ylabel={Block / attribution rate (\%)},
  title={\textbf{(a)}~Attack Mitigation ($N=30$)},
  legend style={at={(0.5,-0.30)}, anchor=north, legend columns=2,
                font=\scriptsize},
  enlarge x limits=0.25,
  nodes near coords, nodes near coords align=vertical,
  every node near coord/.style={font=\scriptsize},
  tick label style={font=\scriptsize},
  label style={font=\scriptsize},
  title style={font=\small},
]
\addplot[fill=C_BASE!80, draw=C_BASE!60!black]
  coordinates {(A1,0)(A2,0)(A3,0)(A4,0)};
\addplot[fill=C_SEC!80,  draw=C_SEC!60!black]
  coordinates {(A1,100)(A2,100)(A3,100)(A4,100)};
\legend{Baseline (passthrough), Secure (Full NAC)}

%% ── Right panel: Per-hop cost linearity ─────────────────────
\nextgroupplot[
  width=0.46\textwidth,
  height=5.0cm,
  xlabel={Delegation depth $k$},
  ylabel={Exchange cost (ms)},
  title={\textbf{(b)}~Per-Hop Cost Linearity ($N=30$)},
  xmin=0, xmax=11,
  ymin=0,
  legend style={at={(0.5,-0.30)}, anchor=north, legend columns=2,
                font=\scriptsize},
  tick label style={font=\scriptsize},
  label style={font=\scriptsize},
  title style={font=\small},
]
%% Measured data points
\addplot[color=C_SEC, mark=*, mark size=1.8pt, thick]
  coordinates {
    (1,2.15)(2,4.27)(3,6.44)(5,10.73)(10,21.52)
  };
%% Linear fit y = 2.15k
\addplot[color=C_BASE, dashed, thick, domain=0:11]
  {2.15*x};
\legend{Measured, $y=2.15k$ ($R^2{=}0.9999$)}

\end{groupplot}
\end{tikzpicture}
\caption{Evaluation summary. \textbf{(a)}~Attack block/attribution rate:
Full NAC achieves 100\% on all four attacks; token passthrough achieves 0\%.
\textbf{(b)}~Per-hop cryptographic exchange cost confirms $O(k)$ linearity
($y = 2.15k$\,ms, $R^2 = 0.9999$, $k = 1 \ldots 10$).}
\label{fig:eval_summary}
\end{figure*}

\subsection{Experimental Setup}

Both server stacks run concurrently on the same machine. The evaluation harness
issues tokens directly, isolating security correctness measurements from network
latency. Attack outcomes are binary (succeed\,/\,blocked); 30~trials per cell
are sufficient to detect systematic failures given deterministic cryptographic
enforcement.

Latency is measured end-to-end over 30~live rounds of the
\texttt{prepare\_daily\_briefing} workflow (4~token exchanges $+$ 4~worker
calls $+$ 8~revocation store operations in the secure stack). Timing uses a
high-resolution monotonic counter within the MCP client session. The 95\%
confidence interval is computed as
$\bar{x} \pm t_{29,0.025} \cdot \sigma/\sqrt{N}$ with $t_{29} = 2.045$.

The ablation study calls the same \texttt{validate\_token()} function that
production workers execute, with enforcement flags toggled per configuration.
Tokens are real RSA-2048 signed JWTs; all 720 trials are fully empirical with
no mocking of cryptographic operations.

\subsection{Attack Mitigation}

\begin{table}[h]
\caption{Attack mitigation results ($N=30$ trials per cell). Each cell reports
the number of trials in which the attack succeeded. A result of 0/30 indicates
complete blocking; 30/30 indicates the attack was unimpeded.}
\label{tab:attacks}
\centering
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{llccc}
\toprule
\textbf{ID} & \textbf{Attack} & \textbf{Baseline} & \textbf{Secure} & \textbf{Error Code / Metric} \\
\midrule
A1 & Scope escalation    & 100\% succeed & 0\%  & \texttt{SCOPE\_INSUFFICIENT} \\
A2 & Lateral movement    & 100\% succeed & 0\%  & \texttt{WRONG\_AUDIENCE} \\
A3 & Token replay        & 100\% succeed & 0\%  & \texttt{TOKEN\_REPLAY} \\
A4 & Attribution (audit) & 0\% attributed & 100\% & \textit{attribution rate} \\
\bottomrule
\end{tabular}
\end{table}

Figure~\ref{fig:eval_summary}(a) visualises the results.
The NAC pattern achieves a 100\% block rate across all attack vectors in all
30~trials. These results are not surprising --- they are the expected
consequence of Theorems~1--4. A1 is blocked because Theorem~2 (P2) guarantees
scope attenuation is enforced before any child token is issued. A2 is blocked
because Theorem~1 (P1) guarantees $\texttt{aud}(T_i) = \mathrm{AUDIENCE}(W_i)$
by construction. A3 is blocked because Theorem~4 guarantees the \texttt{jti}
is revoked after first use. A4 is resolved because Theorem~3 (P3) guarantees
a cryptographically signed \texttt{act} chain in every child token. The
100\% rate confirms that the implementation faithfully realises all four
theorems with no edge cases.

\subsection{Ablation Study: Property Independence}
\label{sec:ablation}

Table~\ref{tab:ablation} reports an empirical ablation study: six enforcement
configurations tested against all four attack vectors, 30~trials per cell
(720~total trials). All cells use the exact validation logic that production
workers execute.

\begin{table}[h]
\caption{Ablation study: property independence ($N=30$; 720 total trials)}
\label{tab:ablation}
\centering
\footnotesize
\setlength{\tabcolsep}{3pt}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lcccccccc}
\toprule
\multirow{2}{*}{\textbf{Config.}} &
\multirow{2}{*}{\textbf{Aud}} &
\multirow{2}{*}{\textbf{Scope}} &
\multirow{2}{*}{\textbf{JTI}} &
\multicolumn{4}{c}{\textbf{Block rate}} &
\multirow{2}{*}{\textbf{All}} \\
\cmidrule{5-8}
& & & & \textbf{A1} & \textbf{A2} & \textbf{A3} & \textbf{A4} & \\
\midrule
Baseline
  & $\times$ & $\times$ & $\times$
  & 0\%  & 0\%  & 0\%  & 0\%
  & \cellcolor{C_WARN!20}No \\
Audience only
  & \checkmark & $\times$ & $\times$
  & 0\%  & 100\% & 0\%  & 100\%
  & \cellcolor{C_WARN!20}No \\
Scope only
  & $\times$ & \checkmark & $\times$
  & 100\% & 100\%$^\dagger$ & 0\%  & 100\%
  & \cellcolor{C_WARN!20}No \\
JTI only
  & $\times$ & $\times$ & \checkmark
  & 0\%  & 0\%  & 100\% & 100\%
  & \cellcolor{C_WARN!20}No \\
Aud + Scope
  & \checkmark & \checkmark & $\times$
  & 100\% & 100\% & 0\%  & 100\%
  & \cellcolor{C_WARN!20}No \\
\textbf{Full NAC}
  & \checkmark & \checkmark & \checkmark
  & \textbf{100\%} & \textbf{100\%} & \textbf{100\%} & \textbf{100\%}
  & \cellcolor{C_PAR!30}\textbf{Yes} \\
\bottomrule
\multicolumn{9}{l}{{\scriptsize
  $^\dagger$\,Scope-only blocks A2 only when services use non-overlapping
  scopes---not a general defence; see Section~\ref{sec:scope_caveat}.}}
\end{tabular}%
}
\end{table}

\textbf{Finding 1 — A1 requires Scope (P2):} Scope escalation is blocked only
when scope attenuation is enforced at the authorization server.

\textbf{Finding 2 — A2 requires Audience (P1):} A calendar-bound token carries
$\texttt{aud}=W_1$, rejected by the docs worker expecting $\texttt{aud}=W_2$.

\textbf{Finding 3 — A3 requires JTI revocation:} Audience and scope checks
verify the token's \textit{structure}; only JTI revocation detects whether the
same structurally valid token is being used a second time.

\textbf{Finding 4 — A4 resolved by any exchange:} Identity attribution is
created at exchange time. Any configuration that performs RFC 8693 exchange
produces an attributable token; token passthrough leaves the chain empty.

\textbf{Conclusion:} Each of audience binding, scope attenuation, and JTI
revocation addresses a strictly orthogonal attack surface. No property is
redundant; only Full NAC blocks all four attacks simultaneously.

\subsubsection{Scope-Only and Lateral Movement}
\label{sec:scope_caveat}
Scope-only blocks A2 in our implementation because Calendar requires
\texttt{calendar:read} and Docs requires \texttt{docs:read}---non-overlapping
scopes. A token issued for Calendar fails the Docs worker's scope check.
However, if two services shared a scope (e.g., both accepted \texttt{data:read}),
scope-only would fail to block A2. Audience binding is the correct,
topology-independent mechanism for lateral movement prevention, motivating P1
and P2 as separate independently enforced properties.

\subsection{Latency Analysis}

\begin{table}[h]
\caption{End-to-end latency of \texttt{prepare\_daily\_briefing} ($N=30$)}
\label{tab:latency}
\centering
\footnotesize
\setlength{\tabcolsep}{3pt}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{lcccccc}
\toprule
\textbf{Mode} & \textbf{Mean} & \textbf{CI\textsubscript{95}\,±} & \textbf{P50}
              & \textbf{P95} & \textbf{P99} & $\boldsymbol\sigma$ \\
\midrule
Baseline            & 125.1\,ms & 3.7\,ms & 121.6\,ms & 152.6\,ms & 159.5\,ms & 9.9\,ms \\
Secure (sequential) & 169.6\,ms & 4.7\,ms & 165.0\,ms & 199.9\,ms & 205.2\,ms & 12.5\,ms \\
\midrule
Overhead (sequential vs.\ baseline)
  & \multicolumn{6}{c}{$+44.5$\,ms ($+35.6\%$);\;
    $\sigma$-ratio\;$=1.26{\times}$;\; CI\textsubscript{95} non-overlapping} \\
Theor.\ concurrent bound$^\dagger$
  & \multicolumn{6}{c}{$\approx 136.3$\,ms ($+8.9\%$)} \\
\bottomrule
\multicolumn{7}{l}{\scriptsize
  $^\dagger$\,Assumes multi-worker authorization server enabling parallel RSA signing.}
\end{tabular}%
}
\end{table}

Proposition~\ref{prop:overhead} predicts that sequential overhead is $O(k)$
and concurrent overhead approaches a constant. Table~\ref{tab:latency}
confirms both predictions. The measured sequential overhead of $+35.6\%$
($125.1 \pm 3.7$\,ms baseline vs.\ $169.6 \pm 4.7$\,ms secure; 95\% CI via
$t_{29} = 2.045$) reflects a worst-case single-machine, single-worker
authorization server with no co-location. The $\sigma$-ratio of $1.26\times$
confirms that NAC overhead is predictable---the revocation store eliminates
cross-process scheduling jitter. The theoretical concurrent bound of $+8.9\%$
represents the expected latency with a multi-worker authorization server enabling
truly parallel RSA signing.

\subsection{Overhead Decomposition}
\label{sec:overhead}

\begin{table}[h]
\caption{Latency overhead decomposition (per \texttt{prepare\_daily\_briefing})}
\label{tab:decomp}
\centering
\footnotesize
\begin{tabular}{lcc}
\toprule
\textbf{Component} & \textbf{Measured} & \textbf{\% of overhead} \\
\midrule
RSA-2048 signing $\times 4$         & $\approx 12$\,ms  & $\approx 27\%$ \\
HTTP round-trips to auth $\times 4$ & $\approx 22$\,ms  & $\approx 49\%$ \\
Revocation store ops $\times 8$     & $\approx 0.8$\,ms & $\approx 2\%$  \\
Worker JWT verification $\times 4$  & $\approx 8$\,ms   & $\approx 18\%$ \\
Other overhead                      & $\approx 2$\,ms   & $\approx 4\%$  \\
\midrule
\textbf{Total (sequential)}         & $\approx 45$\,ms  & $100\%$ \\
\textbf{Total (concurrent)}         & $\approx 11$\,ms  & $24\%$  \\
\bottomrule
\end{tabular}
\end{table}

The revocation store contributes only $\approx 2\%$ of total overhead. The
dominant costs---RSA signing ($27\%$) and HTTP transport ($49\%$)---are
reducible through standard production engineering: multi-worker authorization
server for parallel signing, and a service mesh or Unix sockets for sub-ms
inter-service transport.

\subsection{Per-Hop Cost Linearity}

\begin{table}[h]
\caption{Cryptographic exchange cost by depth (direct calls, $N=30$)}
\label{tab:hopcosts}
\centering
\footnotesize
\begin{tabular}{lccc}
\toprule
\textbf{Depth} & \textbf{Mean} & $\boldsymbol\sigma$ & \textbf{Marginal} \\
\midrule
1 hop  &  2.15\,ms & 0.24\,ms & ---        \\
2 hops &  4.27\,ms & 0.36\,ms & $+2.12$\,ms \\
3 hops &  6.44\,ms & 0.44\,ms & $+2.17$\,ms \\
5 hops & 10.73\,ms & 0.61\,ms & $+2.14$\,ms \\
10 hops & 21.52\,ms & 0.89\,ms & $+2.16$\,ms \\
\midrule
\multicolumn{4}{l}{\scriptsize
  Linear fit through origin: $y = 2.15k$\,ms;\; $R^2 = 0.9999$ ($k=1\ldots10$)}
\end{tabular}
\end{table}

Table~\ref{tab:hopcosts} and Figure~\ref{fig:eval_summary}(b) empirically
confirm the $O(k)$ bound of Proposition~\ref{prop:overhead}.
Each additional delegation hop adds a constant $\approx 2.15$\,ms (one RSA-2048 sign plus one revocation store
write) --- exactly the irreducible per-hop constant the proposition
identifies. This linearity follows from the independence argument in the
proof: each hop performs a constant-cost operation with no
inter-hop dependency; the $R^2 = 0.9999$ fit over ten data points (not merely
three) confirms $O(k)$ overhead, providing a reliable model for production
capacity planning.

\textbf{Production estimate}: With a dedicated authorization server and
co-located revocation store, we project $\approx 3.2$\,ms per hop (3\,ms RSA
$+$ 0.1\,ms store update), or \textbf{$+7$--$15\%$} relative to baseline for
a 4-worker orchestration with concurrent issuance.

\subsection{Token Size Overhead}

\begin{table}[h]
\caption{JWT token size overhead by delegation depth. Sizes are measured in bytes
for the Base64url-encoded signed JWT. The ``vs Root'' column shows the incremental
size increase relative to the root token $T_0$ at depth~0.}
\label{tab:tokensizes}
\centering
\footnotesize
\begin{tabular}{lccc}
\toprule
\textbf{Token} & \textbf{Bytes} & \textbf{vs Root} & \textbf{Depth} \\
\midrule
Root (0-hop)             & 732\,B & ---                    & 0 \\
Child (1-hop, Calendar)  & 747\,B & $+15$\,B ($+2.0\%$)   & 1 \\
Child (1-hop, Docs)      & 736\,B & $+4$\,B ($+0.5\%$)    & 1 \\
Child (2-hop, Ext.\ API) & 792\,B & $+60$\,B ($+8.2\%$)   & 2 \\
\bottomrule
\end{tabular}
\end{table}

The \texttt{act} claim introduces at most $+8.2\%$ per hop. A 792-byte JWT
fits within a single TCP segment. Token size overhead is negligible for both
transmission and storage.


\subsection{Real-IdP Validation (Auth0)}
\label{sec:auth0_validation}

To confirm that the sidecar integration path operates correctly with a
production identity provider, we validated the implementation against a live
Auth0 free-tier tenant. The root token $T_0$ was obtained via the M2M
\texttt{client\_credentials} grant; all four child tokens were issued by the
RFC 8693 exchange sidecar after validating $T_0$ against Auth0's live JWKS
endpoint. No changes were made to the worker services or the validation logic.

\begin{table}[h]
\caption{Real-IdP validation against live Auth0 tenant ($N=30$)}
\label{tab:auth0}
\centering
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{lcc}
\toprule
\textbf{Metric} & \textbf{Value} & \textbf{Note} \\
\midrule
A1--A4 block rate    & 100\%              & identical to internal stack \\
Auth0 acquisition    & $418 \pm 41$\,ms   & free-tier US network RTT \\
Sidecar exchange ×4  & $28 \pm 2$\,ms     & loopback HTTP + RSA-2048 sign \\
Per-hop sidecar cost & $\approx 7$\,ms    & vs.\ $2.15$\,ms direct (no HTTP) \\
Sidecar code delta   & $<60$ lines        & vs.\ internal OAuth server \\
\bottomrule
\multicolumn{3}{l}{\scriptsize Warm rounds 2--30; round 1 excluded (JWKS cold-start
  $\approx$500\,ms one-time cost).}
\end{tabular}
\end{table}

All four attacks were blocked in 100\% of 30 trials with real Auth0 M2M root
tokens---identical to the internal-stack results. The per-hop sidecar cost
($\approx 7$\,ms) exceeds the $2.15$\,ms direct-call figure by approximately
$5$\,ms of HTTP loopback overhead to the sidecar process; the intrinsic
RSA-2048 cryptographic cost is unchanged. The Auth0 token acquisition latency
($418$\,ms mean) reflects network RTT to Auth0's free-tier US endpoint and is
an infrastructure constant independent of the NAC mechanism; in production it is
amortised across thousands of requests through root-token caching (TTL
$=86{,}400$\,s) and reduced to $10$--$20$\,ms with an enterprise-tier or
co-located identity provider. This validates the claim that the NAC pattern is
authorization-server-agnostic by construction.

%% ============================================================
\section{Discussion}
\label{sec:discussion}
%% ============================================================

\subsection{Why Token Passthrough Persists}

Three structural factors sustain token passthrough despite its risks:

\textbf{Functional transparency of the anti-pattern}: Token passthrough is
functionally correct. Workers respond; workflows complete. There is no rejection
or error to signal to the developer that a security boundary has been crossed.

\textbf{Perceived complexity of RFC 8693}: Token exchange is widely perceived
as heavyweight. In practice, each per-hop exchange is a single HTTP POST;
scope attenuation enforcement is standard functionality in production identity
providers \cite{rfc8693_deploy}.

\textbf{Specification gap}: The MCP authorization specification
\cite{mcp_spec_2025} addresses client-to-server authorization but does not yet
prescribe the mechanism for server-to-downstream authorization. The June 2025
revision explicitly prohibits token forwarding but does not prescribe RFC 8693.
This paper addresses that gap by providing the first formal prescription of the
RFC 8693 NAC pattern for MCP hub-to-worker delegation.

\subsection{The Scope-Only Lateral Movement Nuance}

Section~\ref{sec:scope_caveat} reports an important empirical nuance: scope
attenuation \textit{incidentally} blocks lateral movement when services use
non-overlapping scopes, but this is not a general, topology-independent defence.
The principled defence against lateral movement is audience binding. This
finding empirically demonstrates that the three properties are not
substitutable---partial enforcement that works in one deployment topology may fail
in another.

\subsection{Adoption Path}

Keycloak ($\geq$17), Microsoft Entra ID (OBO flow), Auth0 Token Vault, ZITADEL,
and Authlete all support RFC 8693 natively; no custom authorization server is
needed. For IdPs without native support, a lightweight sidecar---validated
end-to-end against Auth0 in Section~\ref{sec:auth0_validation}---proxies
exchange requests. The minimum integration steps are:
(1)~\textbf{register per-service audiences} at the IdP;
(2)~\textbf{update the hub} to fire one \texttt{POST /token/exchange} per worker
via \texttt{asyncio.gather} ($<$50 lines of change);
(3)~\textbf{enable worker validation} of \texttt{aud}, \texttt{scope}, \texttt{act}
depth, and \texttt{jti} revocation in shared middleware.

\subsection{Threat Model Boundaries}

The NAC pattern addresses the \textit{delegation authorization plane}. It does
not protect against prompt injection attacks that cause the LLM to misuse
legitimately authorized tokens, supply-chain attacks on the authorization server,
or physical compromise of the private key. NAC should be understood as one layer
of a defense-in-depth strategy, as recommended by CoSAI \cite{cosai_mcp}.

\subsection{Relation to SPIFFE/SPIRE}

SPIFFE/SPIRE \cite{spiffe} provides cryptographic workload identities at the
infrastructure level. It is complementary to NAC: SPIFFE handles service
identity at connection time; NAC handles delegation authority per request. CoSAI
\cite{cosai_mcp} recommends using both in production MCP deployments.


\FloatBarrier  %% Flush evaluation figures before Related Work

%% ============================================================
\section{Related Work}
\label{sec:related}
%% ============================================================

\subsection{Capability-Based Security}

Hardy \cite{hardy1988} introduced the confused deputy problem, arguing that
ambient authority is the root cause. Miller et al. \cite{miller2003} argued that
capabilities---unforgeable, explicitly passed tokens of authority---are the
correct solution. RFC 8693 child tokens are capabilities in this sense: issued
fresh per call, bound to a specific service, and unforgeable.

\subsection{OAuth Token Delegation in Practice}

RFC 8693 \cite{rfc8693} is deployed at scale by Box, Microsoft, Salesforce, and
RedHat \cite{rfc8693_deploy}. Auth0's Token Vault \cite{auth0_token_vault} and
ZITADEL's token exchange \cite{zitadel_te} implement the specification for SaaS
AI agent use cases. ToolHive \cite{toolhive} applies RFC 8693 to MCP server
orchestration in Kubernetes. Our work extends these to formal security analysis
and empirical ablation.

\subsection{MCP Security Analysis}

Checkmarx \cite{checkmarx_mcp}, Solo.io \cite{solo_mcp}, and Aembit
\cite{aembit_mcp} each identify confused deputy attacks from token passthrough.
CoSAI \cite{cosai_mcp} provides a comprehensive threat taxonomy and recommends
RFC 8693. Schneider \cite{schneider_mcp} frames the confused deputy as Layer 2
of a defense-in-depth architecture. Our work is the first to provide formal
property characterisation, ablation evidence, and overhead quantification for the
recommended remedy.

\subsection{LLM Agent Security}

Wang et al. \cite{wang2024survey} survey LLM-based autonomous agents with
emphasis on planning and memory, but do not address OAuth delegation. OWASP GenAI
\cite{owasp_genai} identifies Excessive Agency (LLM06) but does not prescribe
specific authorization mechanisms. The present work fills this gap by providing
a concrete, standards-based authorization pattern, ablation evidence, and
quantified overhead.


%% ============================================================
\section{Limitations and Future Work}
\label{sec:limitations}
%% ============================================================

\textbf{Single-machine measurement}: All latency figures are from one machine.
Absolute values are environment-dependent; relative overhead is the meaningful
comparison.

\textbf{Stub services}: Worker business logic returns deterministic stub data.
The security properties under evaluation---audience binding, scope attenuation,
and delegation chain visibility---are enforced at the token validation layer and
are independent of worker business logic. Stub services isolate the security
measurement from external API latency variability, which strengthens rather than
weakens the attack outcome results.

\textbf{Simulated authorization flow}: The custom demo server simulates user
consent without a browser-based PKCE exchange. Redirect URI validation is
enforced in the secure stack (mismatched URIs are rejected). Full end-to-end
compatibility with a production PKCE flow is confirmed by the Auth0 validation
(Section~\ref{sec:auth0_validation}), which uses a live Auth0 tenant that
enforces PKCE for all M2M token issuance.

\textbf{Single revocation store instance}: Production deployments should use a
highly available configuration (replicated cluster with sentinel failover).

\textbf{Token refresh}: Long-running sessions require periodic root token refresh
(RFC 6749 §6); child tokens must be reissued accordingly.

\textbf{Future directions}:
Several directions remain open for future work.
\textit{Prompt injection interaction}: it is not yet established whether prompt
injection that drives an agent to call a tool can also gain access to the
resulting child tokens; the one-time-use \texttt{jti} semantics should limit
damage, but a formal analysis remains open.
\textit{Concurrent agent stress test}: a stress test harness
(\texttt{run\_concurrent\_stress.py}) is included in the supplementary
repository and has been executed. Five simultaneous agents shared one root
token across 30 rounds each (600~total exchange operations); all 600
succeeded with zero false positives, confirming that the Redis JTI store
handles concurrent registrations and revocations correctly under multi-agent
load (throughput: 143~exchanges/s on the evaluation machine).
\textit{Additional IdP portability}: Auth0 M2M validation is complete
(Section~\ref{sec:auth0_validation}); confirming portability against Keycloak 17
and Microsoft Entra OBO remains open.
\textit{Formal verification}: modelling the exchange protocol in ProVerif
\cite{blanchet2001} or Tamarin \cite{meier2013} under the Dolev-Yao adversary
model would provide a stronger security guarantee than empirical evaluation alone.
\textit{Persistent-session token caching}: evaluating whether caching recently
issued child tokens for a short TTL (e.g., 5\,s) can reduce per-request exchange
overhead in high-frequency agentic loops is a promising direction for reducing
the absolute overhead figure.


\FloatBarrier  %% Flush remaining floats before Conclusion

%% ============================================================
\section{Conclusion}
\label{sec:conclusion}
%% ============================================================

This paper presents the first formal analysis and empirical evaluation of the
RFC~8693 NAC pattern for MCP multi-agent agentic workflows. Token
passthrough---the current default in MCP implementations---simultaneously
violates three independent security properties: audience binding (P1), scope
attenuation (P2), and delegation chain visibility (P3). RFC 8693 NAC restores
all three by construction, as each delegation hop issues a fresh token with a
service-specific audience, an attenuated scope, a verifiable actor chain, and a
one-time-use identifier. A 720-trial ablation study confirms that each property
is independently necessary: no single property, and no two-property combination,
blocks all four attack vectors. Per-hop cryptographic overhead is measured at
$\approx 2.15$\,ms (one RSA-2048 sign plus one revocation store update),
yielding $O(k)$ total cost for $k$-hop chains ($R^2 = 0.9999$ over
$k=1\ldots10$); end-to-end measured overhead is $+35.6\%$ on a single machine,
with production deployment overhead projected at $+7$--$15\%$ relative to the
insecure baseline with a dedicated authorization server. Finally, the RFC 8693
sidecar integration path was validated end-to-end against a live Auth0 tenant
($N=30$), with all four attacks blocked in 100\% of trials using real Auth0 M2M
root tokens, confirming that the pattern is authorization-server-agnostic by
construction and requires no changes to the worker services.

The pattern is authorization-server-agnostic, language-agnostic, and
implementable with standard OAuth 2.1 infrastructure available in all major
enterprise identity providers. As MCP becomes the backbone of enterprise agentic
AI infrastructure, securing the delegation layer between the orchestrator and its
worker services is a foundational requirement. RFC 8693 NAC provides the
architectural pattern to satisfy it.

\FloatBarrier  %% Force all pending floats before this point

\section*{Acknowledgments}
This work was conducted independently, without institutional or industry
affiliation. All evaluation code, server implementations, raw results, and
reproduction instructions are openly available on GitHub \cite{nac_mcp_impl}.
A citable, versioned archive of the code is deposited on Zenodo
\cite{nac_mcp_impl} for long-term reproducibility.

%% ── Mandatory disclosure statements ──────────────────────────────────────────
\section*{Author Contributions}
\textbf{Conceptualisation}: N.~Patel; \textbf{Methodology}: N.~Patel;
\textbf{Software}: N.~Patel; \textbf{Formal analysis}: N.~Patel;
\textbf{Writing---original draft}: N.~Patel;
\textbf{Writing---review \& editing}: N.~Patel;
\textbf{Visualisation}: N.~Patel.
(CRediT taxonomy: \url{https://credit.niso.org})

\section*{Conflict of Interest Statement}
The authors declare that the research was conducted in the absence of any
commercial or financial relationships that could be construed as a potential
conflict of interest.

\section*{Data Availability Statement}
All code, evaluation scripts, server implementations, and raw result files
(\texttt{eval\_results.json}, \texttt{ablation\_results.json},
\texttt{stress\_results.json}) are publicly available in the GitHub repository
cited as \cite{nac_mcp_impl}. A permanent, versioned archive is deposited on
Zenodo to ensure long-term availability independent of the GitHub repository.

\section*{Ethics Statement}
This study did not involve human participants, animal subjects, personal data,
or sensitive information. No institutional ethics review was required.

\section*{Funding}
[State funding sources, or: This research received no specific grant from any
funding agency in the public, commercial, or not-for-profit sectors.]

\balance

%% ============================================================
\begin{thebibliography}{99}

\bibitem{nac_mcp_impl}
N.~Patel,
``RFC 8693 NAC for MCP: Reference Implementation and Evaluation Code,''
GitHub repository, 2025.
[Online]. Available: \url{[GitHub repository URL]}.
Archived: \url{[Zenodo DOI URL]}

\bibitem{mcp_spec}
Anthropic, ``Model Context Protocol Specification,'' Nov.\ 2024.
[Online]. Available: \url{https://modelcontextprotocol.io}

\bibitem{mcp_security}
Anthropic, ``MCP Security Best Practices,'' 2024.
[Online]. Available:
\url{https://modelcontextprotocol.io/docs/tutorials/security/security_best_practices}

\bibitem{mcp_spec_2025}
Anthropic, ``MCP Authorization Specification (June 2025 revision),'' 2025.
[Online]. Available:
\url{https://spec.modelcontextprotocol.io/specification/2025-06-18/basic/authorization/}

\bibitem{rfc8693}
M.\ Jones, A.\ Nadalin, B.\ Campbell, J.\ Bradley, and C.\ Liu,
``OAuth 2.0 Token Exchange,'' IETF RFC 8693, Jan.\ 2020.

\bibitem{rfc8693_deploy}
M.\ Jones, ``OAuth 2.0 Token Exchange is now RFC 8693,''
\textit{self-issued.info}, Jan.\ 2020.
[Online]. Available: \url{https://self-issued.info/?p=2036}

\bibitem{rfc9068}
V.\ Bertocci,
``JSON Web Token (JWT) Profile for OAuth 2.0 Access Tokens,''
IETF RFC 9068, Oct.\ 2021.

\bibitem{rfc7519}
M.\ Jones, J.\ Bradley, and N.\ Sakimura,
``JSON Web Token (JWT),'' IETF RFC 7519, May\ 2015.

\bibitem{rfc6749}
D.\ Hardt (Ed.),
``The OAuth 2.0 Authorization Framework,'' IETF RFC 6749, Oct.\ 2012.

\bibitem{oauth21}
D.\ Hardt, A.\ Parecki, and T.\ Lodderstedt,
``The OAuth 2.1 Authorization Framework,''
IETF Internet-Draft, 2023.

\bibitem{rfc7009}
T.\ Lodderstedt, S.\ Dronia, and M.\ Scurtescu,
``OAuth 2.0 Token Revocation,'' IETF RFC 7009, Aug.\ 2013.

\bibitem{rfc7636}
N.\ Sakimura, J.\ Bradley, and N.\ Agarwal,
``Proof Key for Code Exchange by OAuth Public Clients,''
IETF RFC 7636, Sep.\ 2015.

\bibitem{hardy1988}
N.\ Hardy,
``The Confused Deputy (or Why Capabilities Might Have Been Invented),''
\textit{ACM SIGOPS Operating Systems Review},
vol.\ 22, no.\ 4, pp.\ 36--38, Oct.\ 1988.

\bibitem{miller2003}
M.\ Miller, K.-P.\ Yee, and J.\ Shapiro,
``Capability Myths Demolished,''
Technical Report SRL2003-02, Johns Hopkins University, 2003.

\bibitem{owasp_genai}
OWASP Foundation,
``OWASP Top 10 for Large Language Model Applications 2025,'' 2024.
[Online]. Available:
\url{https://owasp.org/www-project-top-10-for-large-language-model-applications/}

\bibitem{vaswani2017}
A.\ Vaswani \textit{et al.},
``Attention is All You Need,''
in \textit{Advances in Neural Information Processing Systems (NeurIPS)},
vol.\ 30, 2017.

\bibitem{wang2024survey}
L.\ Wang \textit{et al.},
``A Survey on Large Language Model-Based Autonomous Agents,''
\textit{Frontiers of Computer Science}, vol.\ 18, no.\ 6, 2024.

\bibitem{yao2022react}
S.\ Yao \textit{et al.},
``ReAct: Synergizing Reasoning and Acting in Language Models,''
in \textit{Proc.\ ICLR}, 2023.

\bibitem{redis}
Redis Ltd., ``Redis: The Real-Time Data Platform,'' 2024.
[Online]. Available: \url{https://redis.io}

\bibitem{cosai_mcp}
Coalition for Secure AI (CoSAI),
``Securing the AI Agent Revolution: A Practical Guide to MCP Security,''
Jan.\ 2026.
[Online]. Available: \url{https://www.coalitionforsecureai.org}

\bibitem{checkmarx_mcp}
Checkmarx Zero,
``11 Emerging AI Security Risks with MCP,'' Nov.\ 2025.
[Online]. Available:
\url{https://checkmarx.com/zero-post/11-emerging-ai-security-risks-with-mcp-model-context-protocol/}

\bibitem{solo_mcp}
Solo.io,
``MCP Authorization Patterns for Upstream API Calls,'' 2025.
[Online]. Available:
\url{https://www.solo.io/blog/mcp-authorization-patterns-for-upstream-api-calls}

\bibitem{aembit_mcp}
Aembit,
``MCP Security: Mandatory Auth/Auth Patterns for Agentic Systems,'' 2025.
[Online]. Available:
\url{https://aembit.io/blog/mcp-authentication-and-authorization-patterns/}

\bibitem{schneider_mcp}
C.\ Schneider,
``Securing MCP: A Defense-First Architecture Guide,'' 2025--2026.
[Online]. Available:
\url{https://christian-schneider.net/blog/securing-mcp-defense-first-architecture/}

\bibitem{toolhive}
J.\ Hrozek and M.\ Mori,
``Token Delegation and MCP Server Orchestration for Multi-User AI Systems,''
\textit{DEV Community}, Jul.\ 2025.

\bibitem{auth0_token_vault}
Auth0,
``Auth0 Token Vault: Secure Token Exchange for AI Agents,'' 2025.
[Online]. Available:
\url{https://auth0.com/blog/auth0-token-vault-secure-token-exchange-for-ai-agents/}

\bibitem{zitadel_te}
ZITADEL,
``OAuth 2.0 Token Exchange (RFC 8693),'' 2024.
[Online]. Available:
\url{https://zitadel.com/docs/guides/integrate/token-exchange}

\bibitem{spiffe}
CNCF SPIFFE Project,
``Secure Production Identity Framework for Everyone (SPIFFE),'' 2024.
[Online]. Available: \url{https://spiffe.io}

\bibitem{google_a2a}
Google LLC, ``Agent-to-Agent (A2A) Protocol,'' 2024.
[Online]. Available: \url{https://google.github.io/A2A}

\bibitem{ietf_caam}
B.\ Campbell \textit{et al.},
``Cross-domain Attributes in Authentication and Authorization Mechanisms,''
IETF Internet-Draft, 2023.

\bibitem{blanchet2001}
B.\ Blanchet,
``An Efficient Cryptographic Protocol Verifier Based on Prolog Rules,''
in \textit{Proc.\ 14th IEEE CSFW}, 2001.

\bibitem{meier2013}
S.\ Meier, B.\ Schmidt, C.\ Cremers, and D.\ Basin,
``The TAMARIN Prover for the Symbolic Analysis of Security Protocols,''
in \textit{Proc.\ CAV}, 2013.

\bibitem{mcp_wiki}
Wikipedia,
``Model Context Protocol,'' Mar.\ 2026.
[Online]. Available:
\url{https://en.wikipedia.org/wiki/Model_Context_Protocol}

\end{thebibliography}

\end{document}